Lightness constancy under fixed geometry

### 2017-12-21 Draft1/draft1.tex
December 21 (Winter Solstice) editing

\subsubsection{Illumination Spectra}
We generate illumination spectra  through random statistical sampling of a dataset of naturally occurring illuminants, the \href{http://colorimaginglab.ugr.es/pages/Data}{Granada daylight measurements} \cite{peyvandi2016colorimetric}.  First, we approximate the spectra in the dataset using principle component analysis (PCA). We then decompose the spectra in terms of the PCA basis vectors corresponding to the 6 largest PCA eigenvalues. Next, we sample random spectra from this low 6 dimensional PCA weight space assuming that the distribution of the basis vector weights is a multivariate gaussian whose mean and covariance matrix match the sample mean and covariance obtained from the daylight dataset. This general method for modeling the statistical structure of daylight was introduced by Brainard \& Freeman \cite{BrainardFreeman}. Details of the process we used are provided below.

@@

	\begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure6/Figure6_d.pdf}
        \caption{xy chromaticity diagram}
        \label{fig:xyDiagram}
    \end{subfigure}
	\begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure6/Figure6_e.pdf}
        \caption{Normalized PCA eigenvalues}
        \label{fig:granadaEV}
    \end{subfigure}
      	\begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure6/Figure6_f.pdf}
        \caption{sRGB rendition of samples from Fig.\ref{fig:randomIlluminant}}
        \label{fig:sRGBIlluminant}
    \end{subfigure}
    \caption{{\bf Generating random illumination spectra:} The illumination spectra are generated by random sampling based on the Granada daylight measurements, assuming a multivariate Gaussian distribution of weights in a 6-dimensional PCA representation. The mean and covariance matrix of the Gaussian are chosen by projecting the dataset along the first six PCA directions and taking the mean and covariance matrix of the resulting basis vector weights. (a) The 2600 spectra from the Granada natural daylight dataset. (b) Rescaled version of the Granada dataset. Each spectrum is rescaled so that it has a mean value (over wavelength) of one illumination. (c) Randomly samples generated using the rescaled spectra from Fig.~\ref{fig:rescaledGranada}. (d) CIE xy chromaticity plot of the rescaled dataset (Fig.~\ref{fig:rescaledGranada}) and randomly generated samples from Fig.~\ref{fig:randomIlluminant}. (e) Normalized PCA eigenvalues and cumulative sum of the normalized eigenvalues. The eigenvalues have been normalized by the sum over all eigenvalues. The dotted line at 0.9 on the y-axis is for reference. We used the first six eigenvectors in our analysis. (f) sRGB renditions of the random illuminant samples shown in \ref{fig:randomIlluminant}.}\label{fig:illluminationGeneration}
\end{figure}

@@

The rescaled spectra $I_i^s(\lambda)$ were then mean centered for PCA by subtracting out the mean 
rescaled spectrum, $\bar{I}^{s}(\lambda)$. That mean was obtained
by taking the sample mean over all rescaled spectra in the dataset, $\bar{I}^{s}(\lambda) = \frac{1}{M}\sum_i{I_i^s(\lambda)}$. 
The resulting mean subtracted dataset, $I_i^{MC}(\lambda) = I_{s}(\lambda) - \bar{I}_{s}(\lambda)$
was decomposed as $I_i^{MC}(\lambda) = \sum_j w_{ij}\hat{{\bf e}}_j^{PCA}$, 
where the $\hat{{\bf e}}_j^{PCA}$ are the PCA basis vectors obtained using the
singular value decomposition (SVD) applied to the $I_i^{MC}(\lambda)$ and
the $w_{ij}$ are the weights obtained by projecting each of $I_i^{MC}(\lambda)$ onto the  $\hat{{\bf e}}_j^{PCA}$ $\left( w_{ij} = I_i^{\rm MC}(\lambda)\cdot {\bf \hat{e}}_j^{\rm PCA}\right)$.
We used the basis vectors corresponding to
the largest six SVD eigenvalues, so that $\{j \in [16]\}$.
For the rescaled Granada dataset, these six  eigenvalues account for more than $90\%$ of the variance.
These steps
can be summarized as follows:
\begin{align}
I^G_i(\lambda) \rightarrow I_i^s(\lambda) = \frac{I^G_i(\lambda)}{\int d\lambda I^G_i(\lambda)} \rightarrow I_i^{MC}(\lambda) = I_{s}(\lambda) - \bar{I}_{s}(\lambda) \rightarrow I_i^{MC}(\lambda) \approx \sum_{j = 1}^6 w_{ij}\hat{{\bf e}}_j^{PCA}.
\end{align}

To generate random illuminant spectra $\tilde{I}_i(\lambda)$, we generate weights $\tilde{w}_{ij}$ drawn from random 
the multivariate 
Gaussian distribution with mean $\bar{w}_j = \frac{1}{M}\sum_i w_{ij}$, 
and co-variance $\Sigma_{jj'} = \frac{1}{M} \sum_i \left(w_{ij} -\bar{w}_j\right)\left(w_{ij'} -\bar{w}_{j'}\right) $.
From these weights, the corresponding spectrum $\left( \sum_j \tilde{w}_{ij} \hat{{\bf e}}_j^{PCA} +  \bar{I}_{s} (\lambda)\right)$ is generated.
This spectrum can sometimes have values that are less than zero.  In such cases, the weights are discarded and a new draw obtained, until
the condition $\left( \sum_j \tilde{w}_{ij} \hat{{\bf e}}_j^{PCA} +  \bar{I}_{s} (\lambda)\right) > 0$, for all $\lambda$, is satisifed.
The random illuminant spectrum is then given as
\begin{align}
\tilde{I}(\lambda) = \left( \tilde{I}_{MC}(\lambda) + \bar{I}_{s}(\lambda)\right).
\end{align}
To get a scaling similar to the original Granada spectrum set, we multiply the randomly sampled spectrum with a random number generated uniformly from the interval [0, $\bar{I}_{\rm max}$], where $\bar{I}_{\rm max}$ is the maximum of the 2600 mean values   obtained from the Granada spectra. Finally, to model the fact that illumination spectra can vary in overall intensity, we draw a random number distributed logarithmically in the range [$I_{min}$ and $I_{max}$] and scale the generated spectrum by this number.
In most of our calculations, we set $I_{min} = 1$ and $I_{max} = 300$.

\subsubsection{Reflectance Spectra}

% Figure 7
\begin{figure}
\centering
	\begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure7/Figure7_a.pdf}
        \caption{Natural surface reflectance spectra}
        \label{fig:naturalSurface}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure7/Figure7_b.pdf}
        \caption{Randomly generated samples}
        \label{fig:randomSurface}
    \end{subfigure}
    \begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure7/Figure7_c.pdf}
        \caption{sRGB rendition of samples from \ref{fig:randomSurface}}
        \label{fig:sRGBSurface}
    \end{subfigure}
    \caption{{\bf Generating random surface reflectance spectra:} The surface reflectance spectra are generated by random sampling of natural surface reflectance spectra, similar to the principles used to generate random illuminant spectra. (a) The 632 surface reflectance spectra from the Munsell and Vrhel dataset. (b) Randomly generated samples of reflectance spectra. (c) The sRGB renditions of the random samples generated in \ref{fig:randomSurface}.}\label{fig:surfaceReflectanceGeneration}
\end{figure}

We generated random reflectance spectra using the same principles we applied to generate random illuminant spectra.
We used the Munsell \cite{kelly1943tristimulus} and Vrhel \cite{vrhel1994measurement} surface reflectance 
datasets for sampling the reflectance. 
Let us denote the reflectance dataset as $R_i(\lambda)$, where $\{i \in [1,M]\}$ 
and $M$ is the total number of spectra in the dataset. 
We calculated the mean reflectance spectrum,
$\bar{R}(\lambda)$, by taking the sample mean over all spectra in the dataset, i.e.,
$\bar{R}(\lambda) = \frac{1}{M} \sum_{i=1}^M R_i(\lambda)$. Then, we mean centered 
the reflectance dataset by subtracting the mean spectrum, $R_i^{\rm MC}(\lambda) =  R_i(\lambda)-\bar{R}(\lambda)$. 
We decomposed this mean centered dataset as $R_i^{\rm MC}(\lambda) = \sum_j{w_{ij} \; {\bf \hat{e}}_j^{\rm PCA}}$, where
${\bf \hat{e}}_j^{\rm PCA}$s are PCA basis vectors obtained using SVD applied to $R_i^{\rm MC}(\lambda)$
and $w_{ij}$ are the weights obtained by projecting each of $R_i^{\rm MC}(\lambda)$ onto the vectors ${\bf \hat{e}}_j^{\rm PCA}$ 
$\left( w_{ij} = R_i^{\rm MC}(\lambda)\cdot {\bf \hat{e}}_j^{\rm PCA}\right)$. 
We used the basis vectors corresponding to the largest six SVD eigenvalues. 
For the combined Munsell and Vrhel datasets, these six eigenvalues account for 
more than $90\%$ of the variance. 
To generate random reflectance spectra, we generate samples of weights ($\tilde{w}_{ij}$) drawn from the multivariate Gaussian 
distribution with mean $\bar{w}_j = \frac{1}{M}\sum_i w_{ij}$, 
and co-variance $\Sigma_{jj'} = \frac{1}{M} \sum_i \left(w_{ij} -\bar{w}_j\right)\left(w_{ij'} -\bar{w}_{j'}\right) $. If the randomly sampled weights satisfy the condition $\left( 0 < \sum_j \tilde{w}_{ij}{\bf \hat{e}}_j^{\rm PCA} + \bar{R}(\lambda)<1\right) $ at every $\lambda$, we use them to give the random reflectance spectrum as: $\tilde{R}_i(\lambda) =\sum_j \tilde{w}_{ij}{\bf \hat{e}}_j^{\rm PCA} + \bar{R}(\lambda)$.

@@

For generating the target object reflectance at a particular luminance $(Y_{\rm T})$, the values in the spectrum are 
rescaled by a scalar such that the perceived lightness of the 
target surface under standard daylight illuminant $(D_{65}(\lambda))$ 
equals a designated value. The scaling equals $\frac{Y_{\rm T}}{\int d\lambda \tilde{R}(\lambda) D_{65}(\lambda) \bar{y}(\lambda)}$, with $\bar{y}(\lambda)$ being the CIE luminosity (or luminous efficiency) function. 
$\bar{y}(\lambda)$ describes the average spectral sensitivity of human visual 
perception of brightness. The target reflectance is given by: $\tilde{R}^{\rm T}_i(\lambda) =\tilde{R}_i(\lambda) \cdot\left(\frac{Y_{\rm T}}{\int d\lambda \tilde{R}(\lambda) D_{65}(\lambda) \bar{y}(\lambda)}\right)$.

% NEED TO DECIDE PRECISELY WHERE TO USE THIS PROSE AND  FIGURE
%Figure~\ref{sceneWithCroppedImage} illustrates the steps in this process for one scene.  We start with how we generate a single labeled set of cone responses
%We call our image generation software Virtual world color constancy (VWCC). The details of the software are provided in the supplementary information section. To generate an image, we first select a 3D scene, a target object and a light source from the set of predefined scenes and objects in VWCC. The target object and the light source are inserted in the 3D scene. Fig.\ref{fig:3DScene} shows a 2D sRGB rendition of the 3D scene used in our analysis. The green sphere in the middle is the target object. The light source is not in the camera field of view. Next, we assign a surface reflectance spectrum and an illumination spectrum to each object and light source in the scene. For the target object the surface reflectance is assigned such that the perceived lightness of the target object under D65 daylight spectrum has a fixed value. Now we render a $320\times 240$ pixel$^2$ 2d multispectral images of the scene at 31 frequencies (equal linear spacing between 400 nm - 700 nm) using a graphics rendering system called Mitsuba \cite{jakob2015mitsuba}. For each of the three case, we generate hundreds of such scenes and corresponding multispectral images at 10 linearly spaced lightness levels between 0.2 and 0.6 (see below). We crop a 51x51 pixel$^2$ part of the multispectral image containing the target object for further analysis (\ref{fig:croppedImage}). Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 lightness levels for the three cases considered in this work.

% Figure 9: Methods
\begin{figure}
\centering
\begin{subfigure}[b]{0.25 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_a.png}
        \caption{A typical 3D Scene}
        \label{fig:3DScene}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.19 \textwidth}   
    \hspace{0.1 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_b.png}
        \caption{Cropped Image}
        \label{fig:croppedImage}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.19 \textwidth}
    \hspace{0.1 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_c.png}
        \caption{Optical Image}
        \label{fig:croppedImageWithMosaic}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.2 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_d.png}
        \caption{LMS cone contrast}
        \label{fig:coneContrast}
    \end{subfigure}
    \label{fig:sceneWithCroppedImage}
    \caption{{\bf Generating labeled dataset for computational analysis:}  For the supervised approach employed in this work, the labeled database is generated as follows: (a) A 3D virtual scene, a target object and a light source is selected from the VWCC library. The target object is placed in the camera field of view and assigned a reflectance spectrum with a specific lightness. After assigning spectra to other objects and lights sources in the scene, a multispectral image of the scene is rendered using Mitsuba graphics rendering software. (b) A part of the image, containing the target object, is cropped from the image. (c) The response of the retinal cones to the cropped image is simulated using a model of the visual system. Fig.\ref{fig:croppedImageWithMosaic} shows the optical image incident on the cone mosaic and the location and the identity of the cones (L cones: Red, M cones: Green, S cones: Blue). See text for the details of the model. (d) The cone response are interpolated to get the response of all the three types at each location (demosaicing). Finally, the demosaic response is contrast normalized (see text).}
\end{figure}

\subsection{Model of Visual System} \label{method:Isetbio}

Finally, these multispectral images are used to simulate retinal cone responses using \href{https://github.com/isetbio}{Isetbio} software.

To generate the retinal response one needs to specify the retinal cone mosaic and the size of the retinal image. The retinal cone mosaic is specified in terms of the total number of cones and the proportion of the long(L), medium(M) and the short(S) cones in the mosaic. With these parameters, the software generates the specified number of LMS cones with random positions on a square grid. The size of the retinal image is specified in terms of the field of view and object distance from the eye.

>>>>

\subsubsection{Illumination Spectra}
We generate illumination spectra  through random statistical sampling of a dataset of naturally occurring illuminants, the \href{http://colorimaginglab.ugr.es/pages/Data}{Granada daylight measurements} \cite{peyvandi2016colorimetric}.  These spectra are shown on a log intensity scale in Fig.~\ref{fig:granadaSpectra}, which illustrates that the overall intensity of natural daylights varies over several orders of magnitude, and in normalized format in Fig.~\ref{fig:rescaledGranada}, which illustrates that the relative spectra also vary considerably.  In our procedure, we first approximate the spectra in the dataset using principle component analysis (PCA). We then decompose the spectra in terms of the PCA basis vectors corresponding to the 6 largest PCA eigenvalues. Next, we sample from this low 6 dimensional PCA weight space assuming that the distribution of the basis vector weights is a multivariate gaussian whose mean and covariance matrix match the sample mean and covariance obtained from the daylight dataset. Finallly, we use the sampled weights to reconstruct sampled spectra. This general method for modeling the statistical structure of daylight was introduced by Brainard \& Freeman \cite{BrainardFreeman}. Details of the procedure we used are provided below.

@@

	\begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure6/Figure6_d.pdf}
        \caption{xy chromaticity diagram}
        \label{fig:xyDiagram}
    \end{subfigure}
	\begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure6/Figure6_e.pdf}
        \caption{Normalized PCA eigenvalues}
        \label{fig:granadaEV}
    \end{subfigure}
      	\begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure6/Figure6_f.pdf}
        \caption{sRGB rendition of samples from Fig.\ref{fig:randomIlluminant}}
        \label{fig:sRGBIlluminant}
    \end{subfigure}
    \caption{{\bf Generating random illumination spectra:} The illumination spectra are generated by random sampling based on the Granada daylight measurements, assuming a multivariate Gaussian distribution of weights in a 6-dimensional PCA representation. The mean and covariance matrix of the Gaussian are chosen by projecting the dataset along the first six PCA directions and taking the mean and covariance matrix of the resulting basis vector weights. (a) The 2600 spectra from the Granada natural daylight dataset, plotted on a log intensity scale in arbitrary units. (b) Rescaled version of the Granada dataset. Each spectrum is rescaled so that it has a mean value (over wavelength) of one illumination. (c) Randomly samples generated using the rescaled spectra from Fig.~\ref{fig:rescaledGranada}. (d) CIE xy chromaticity plot of the rescaled dataset (Fig.~\ref{fig:rescaledGranada}) and randomly generated samples from Fig.~\ref{fig:randomIlluminant}. (e) Normalized PCA eigenvalues and cumulative sum of the normalized eigenvalues. The eigenvalues have been normalized by the sum over all eigenvalues. The dotted line at 0.9 on the y-axis is for reference. We used the first six eigenvectors in our analysis. (f) sRGB renditions of the random illuminant samples shown in \ref{fig:randomIlluminant}.}\label{fig:illluminationGeneration}
\end{figure}

@@

The rescaled spectra $I_i^s(\lambda)$ were then mean centered for PCA by subtracting out the mean 
rescaled spectrum, $\bar{I}^{s}(\lambda)$. That mean was obtained
by taking the sample mean over all rescaled spectra in the dataset, $\bar{I}^{s}(\lambda) = \frac{1}{M}\sum_i{I_i^s(\lambda)}$. 
The resulting mean subtracted dataset, $I_i^{MC}(\lambda) = I_{s}(\lambda) - \bar{I}_{s}(\lambda)$
was decomposed as $I_i^{MC}(\lambda) = \sum_j w_{ij}\hat{{\bf e}}_j^{PCA}$, 
where the $\hat{{\bf e}}_j^{PCA}$ are the PCA basis vectors obtained using the
singular value decomposition (SVD) applied to the $I_i^{MC}(\lambda)$ and
the $w_{ij}$ are the weights obtained by projecting each of $I_i^{MC}(\lambda)$ onto the  $\hat{{\bf e}}_j^{PCA}$ $\left( w_{ij} = I_i^{\rm MC}(\lambda)\cdot {\bf \hat{e}}_j^{\rm PCA}\right)$.
We used the basis vectors corresponding to
the largest six SVD eigenvalues, so that $\{j \in [16]\}$.
For the rescaled Granada dataset, these six  eigenvalues account for more than $90\%$ of the variance (Fig.~\ref{fig:granadaEV}).
These steps
can be summarized as follows:
\begin{align}
I^G_i(\lambda) \rightarrow I_i^s(\lambda) = \frac{I^G_i(\lambda)}{\int d\lambda I^G_i(\lambda)} \rightarrow I_i^{MC}(\lambda) = I_{s}(\lambda) - \bar{I}_{s}(\lambda) \rightarrow I_i^{MC}(\lambda) \approx \sum_{j = 1}^6 w_{ij}\hat{{\bf e}}_j^{PCA}.
\end{align}

To generate random illuminant spectra $\tilde{I}_i(\lambda)$, we generate weights $\tilde{w}_{ij}$ drawn from random 
the multivariate 
Gaussian distribution with mean $\bar{w}_j = \frac{1}{M}\sum_i w_{ij}$, 
and co-variance $\Sigma_{jj'} = \frac{1}{M} \sum_i \left(w_{ij} -\bar{w}_j\right)\left(w_{ij'} -\bar{w}_{j'}\right) $.
From these weights, the corresponding spectrum $\left( \sum_j \tilde{w}_{ij} \hat{{\bf e}}_j^{PCA} +  \bar{I}_{s} (\lambda)\right)$ is generated.
This spectrum can sometimes have values that are less than zero.  In such cases, the weights are discarded and a new draw obtained, until
the condition $\left( \sum_j \tilde{w}_{ij} \hat{{\bf e}}_j^{PCA} +  \bar{I}_{s} (\lambda)\right) > 0$, for all $\lambda$, is satisifed.
The random illuminant spectrum is then given as
\begin{align}
\tilde{I}(\lambda) = \left( \tilde{I}_{MC}(\lambda) + \bar{I}_{s}(\lambda)\right).
\end{align}
Fig.~\ref{fig:randomIlluminant} shows a set of 100 spectra sampled in this manner, and may be compared with Fig.~\ref{fig:rescaledGranada}.  Fig.~\ref{fig:xyDiagram} shows the CIE xy chromaticities of these samples and compares these with the chromaticities in the Granada dataset.  Finally, Fig.~\ref{fig:sRGBIlluminant} provides an sRGB rendering of each spectrum in Fig.~\ref{fig:rescaledGranada}, to provide a visual sense of the chromatic variation in the sampled spectra.

Because our multivariate Gaussian model is based on the rescaled spectra, it does not embody the large variation in overall intensity of natural daylights.
To get a scaling similar to the original Granada spectrum set, we multiply the randomly sampled spectrum with a random number generated uniformly from the interval [0, $\bar{I}_{\rm max}$], where $\bar{I}_{\rm max}$ is the maximum of the 2600 mean values obtained from the Granada spectra. Finally, to model the fact that illumination spectra can vary in overall intensity, we draw a random number distributed logarithmically in the range [$I_{min}$ and $I_{max}$] and scale the generated spectrum by this number.
In most of our calculations, we set $I_{min} = 1$ and $I_{max} = 300$.
% WE NEED TO RETHINK THIS SCALING.  CERTAINLY SHOULD JUST BE ONE STEP.  ONE APPROACH WOULD BE TO DRAW UNIFORMLY FROM THE DISTRIBUTION OF SCALE FACTORS OBTAINED FROM THE 2600 
% GRANADA MEASUREMENTS.

\subsubsection{Surface Reflectance Spectra}

% Figure 7
\begin{figure}
\centering
	\begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure7/Figure7_a.pdf}
        \caption{Natural surface reflectance spectra}
        \label{fig:naturalSurface}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure7/Figure7_b.pdf}
        \caption{Randomly generated spectra}
        \label{fig:randomSurface}
    \end{subfigure}
    \begin{subfigure}[b]{0.3 \textwidth}
    \centering
        \includegraphics[width=\textwidth]{../Figures/Figure7/Figure7_c.pdf}
        \caption{sRGB rendition of samples from \ref{fig:randomSurface}}
        \label{fig:sRGBSurface}
    \end{subfigure}
    \caption{{\bf Generating random surface reflectance spectra:} Surface reflectance spectra are generated by random sampling based on measured surface reflectance spectra, using principles similar to those used to generate random illuminant spectra. (a) The 632 surface reflectance spectra from the Munsell and Vrhel datasets. (b) Randomly generated samples of reflectance spectra. (c) sRGB renderings (under CIE D65) of the random samples generated in \ref{fig:randomSurface}.}\label{fig:surfaceReflectanceGeneration}
\end{figure}

We generated random reflectance spectra using the same principles we applied to generate random illuminant spectra.
We combined the Munsell \cite{kelly1943tristimulus} and Vrhel \cite{vrhel1994measurement} surface reflectance 
measurements to create a reflectance database containing 632 reflectances. 
% SAY SOMETHING ABOUT SPECTRAL SAMPLING
Let us denote the reflectance dataset as $R_i(\lambda)$, where $\{i \in [1,M]\}$ 
and $M$ is the total number of spectra in the dataset. 
We calculated the mean reflectance spectrum,
$\bar{R}(\lambda)$, by taking the sample mean over all spectra in the dataset, i.e.,
$\bar{R}(\lambda) = \frac{1}{M} \sum_{i=1}^M R_i(\lambda)$. Then, we mean centered 
the reflectance dataset by subtracting the mean spectrum, $R_i^{\rm MC}(\lambda) =  R_i(\lambda)-\bar{R}(\lambda)$. 
We decomposed this mean centered dataset as $R_i^{\rm MC}(\lambda) = \sum_j{w_{ij} \; {\bf \hat{e}}_j^{\rm PCA}}$, where
the ${\bf \hat{e}}_j^{\rm PCA}$s are PCA basis vectors obtained using SVD applied to $R_i^{\rm MC}(\lambda)$
and the $w_{ij}$ are the weights obtained by projecting each of $R_i^{\rm MC}(\lambda)$ onto the basis vectors ${\bf \hat{e}}_j^{\rm PCA}$ 
$\left( w_{ij} = R_i^{\rm MC}(\lambda)\cdot {\bf \hat{e}}_j^{\rm PCA}\right)$. 
We used the basis vectors corresponding to the largest six SVD eigenvalues. 
For the combined Munsell and Vrhel datasets, these six eigenvalues account for 
more than $90\%$ of the variance. 
To generate random reflectance spectra, we generate samples of weights ($\tilde{w}_{ij}$) drawn from the multivariate Gaussian 
distribution with mean $\bar{w}_j = \frac{1}{M}\sum_i w_{ij}$, 
and co-variance $\Sigma_{jj'} = \frac{1}{M} \sum_i \left(w_{ij} -\bar{w}_j\right)\left(w_{ij'} -\bar{w}_{j'}\right) $. If the randomly sampled weights satisfy the condition $\left( 0 < \sum_j \tilde{w}_{ij}{\bf \hat{e}}_j^{\rm PCA} + \bar{R}(\lambda)<1\right) $ at every $\lambda$, we use them to give the random reflectance spectrum as: $\tilde{R}_i(\lambda) =\sum_j \tilde{w}_{ij}{\bf \hat{e}}_j^{\rm PCA} + \bar{R}(\lambda)$.  Otherwise the draw is discarded and a new set of weights is drawn.

@@

For generating the target object reflectance at a particular luminance $(Y_{\rm T})$, the values in a generated spectrum were 
rescaled such that the surface luminance (as defined above) had the desired value.
The scaling equals $\frac{Y_{\rm T}}{\int d\lambda \tilde{R}(\lambda) D_{65}(\lambda) \bar{y}(\lambda)}$, with $\bar{y}(\lambda)$ being the CIE photopic luminosity (or luminous efficiency) function. 
$\bar{y}(\lambda)$ describes the average spectral sensitivity of human visual 
perception of brightness. The target reflectance is then given by: $\tilde{R}^{\rm T}_i(\lambda) =\tilde{R}_i(\lambda) \cdot\left(\frac{Y_{\rm T}}{\int d\lambda \tilde{R}(\lambda) D_{65}(\lambda) \bar{y}(\lambda)}\right)$.

% Figure 9: Methods
\begin{figure}
\centering
\begin{subfigure}[b]{0.25 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_a.png}
        \caption{Rendered image of a randomly drawn scene}
        \label{fig:3DScene}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.19 \textwidth}   
    \hspace{0.1 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_b.png}
        \caption{Cropped Image}
        \label{fig:croppedImage}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.19 \textwidth}
    \hspace{0.1 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_c.png}
        \caption{Optical Image}
        \label{fig:croppedImageWithMosaic}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.2 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_d.png}
        \caption{LMS cone contrast}
        \label{fig:coneContrast}
    \end{subfigure}
    \label{fig:sceneWithCroppedImage}
    \caption{{\bf Generating labeled dataset for computational analysis:}  For the supervised approach employed in this work, the labeled database is generated as follows: (a) A 3D virtual scene, containing a target object of known surface luminance is created. The target object is placed in the camera field of view. After assigning spectra to the light sources and other other objects in the scene, a multispectral image of the scene is rendered using the Mitsuba software package. (b) The central portion of the rendered image, which contains the target object, is cropped from the image. (c) The response of retinal cones to the cropped image is simulated using the isetbio software package to model early vision. The software models optical blurring and foveal spatial sampling by the trichromatic cone mosaic. Fig.\ref{fig:croppedImageWithMosaic} shows the retinal image incident on the cone mosaic and the location and the identity of the cones (L cones: Red, M cones: Green, S cones: Blue).  (d) The cone responses are interpolated to estimate the responses of all the three types of cone at each location (demosaicing). Finally, the demosaiced images are contrast normalized separately for each cone type.}
\end{figure}

\subsection{Model of Early Visual System} \label{method:Isetbio}

The sampling and rendering steps described above allow us to generate images from scenes that contain an image of a target object of known surface luminance at the center pixel, and where the illuminant spectra and spectra of the other surfaces in the scene are drawn randomly from models of naturally occurring spectra.  One such image is shown in Fig.~\ref{fig:3DScene}, where the green sphere in the middle is the target object and the light sources are not within the camera field of view. We model the encoding of such images by the early visual system, to generate {\em neural images} that provide an abstracted representation of the information available to the brain for producing perceptual representations of the external world.  It is this information for each image, together with the known surface luminance of the corresponding target object, that we will provide to our supervised learning algorithms.  Our training datasets correspond to % SAY WHAT WE DID ONCE WE FINISH EVERYTHING OFF.

In this work, we are interested in how well lightness constancy may be achieved using reasonably information that is local to the image region containing the target patch.
% MAYBE SAY A LITTLE MORE ABOUT WHY - THIS IS A PRINCIPLED STANCE TAKEN IN MOST OF JOHANNES WORK.  OR MAYBE DEFER THAT TO THE DISCUSSION.
To this end, we crop a 51x51 pixel portion of the full rendreed multispectral image that contains the target object and use this for further analysis (\ref{fig:croppedImage}).
Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 lightness levels for the three cases considered in this paper.

Next, the cropped multispectral images are used to simulate retinal cone responses using \href{https://github.com/isetbio}{Isetbio} software. The key steps we model are optical blurring, including axial chromatic aberration, in the formation of the retinal image, spatial sampling by an interleaved mosaic of L, M and S cones, and the Poisson nature of photopigment isomerization.  
% NEED TO WRITE IN THE KEY PARAMETERS USED. PROBABLY MARIMONT AND WANDELL OPTICS, STOCKMAN-SHARPE LMS CONE SPECTRAL SENSITIVITIES, THE L, M, S CONE RATIO IN THE MOSAIC, RECTANGULAR AND REGULAR CONE PACKING, SOME IMPLIED SPATIAL SCALE, AND SOME OVERALL SCALING THAT DRIVES THE MEAN LUMINANCE OF THE RETINAL IMAGES AND THUS THE AMOUNT OF POISSON NOISE.

% WE GOT TO HERE ON DECEMBER 21.
To generate the retinal response one needs to specify the retinal cone mosaic and the size of the retinal image. The retinal cone mosaic is specified in terms of the total number of cones and the proportion of the long(L), medium(M) and the short(S) cones in the mosaic. With these parameters, the software generates the specified number of LMS cones with random positions on a square grid. The size of the retinal image is specified in terms of the field of view and object distance from the eye.
